New Exact Algorithms for Planar Maximum Covering Location by Ellipses Problems\tnoteref{t1}

### 2020-03-28 defesa/article/article.tex
wip

		\author[1]{Danilo Tedeschi}
		\ead{danilo.tedeschi@usp.br}

		\author[2]{Marina Andretta}
		\ead{andretta@gmail.com}

		\address{Department of Applied Mathematics and Statistics, Institute of Mathematical and Computer Sciences, University of São Paulo, Avenida Trabalhador São-carlense, 400, Centro, 13566-590, São Carlos, SP, Brazil.}
		%\author{John Smith}

		%address{California, United States}

		\begin{abstract}
			%% Text of abstract
			Planar Maximum Covering Location by Ellipses is an optimization problem where one wants to place fixed shape ellipses on the plane to cover demand points
			maximizing a function depending on the value of covered points.
			We propose new exact algorithms for two versions of this problem, one where the ellipses have to be parallel to the coordinate axis, and another where they can be freely rotated. 
			Besides finding optimal solutions for previously published instances, including the ones where no optimal solution was known, both algorithms proposed by us were able to obtain optimal solutions for some new larger instances 	with up to seven hundred demand points and five ellipses.
		\end{abstract}

		\begin{keyword}
			Combinatorial optimization \sep planar maximal covering location problem \sep planar covering by ellipses \sep exact algorithms.
			%% keywords here, in the form: keyword \sep keyword

@@

		\end{keyword}

	\end{frontmatter}

@@

\clearpage
\section*{References}
	%% New version of the num-names style
	\bibliographystyle{elsarticle-num}
	\bibliography{../references}

>>>>

		\author[add1]{Danilo Tedeschi}
		\ead{danilo.tedeschi@usp.br}

		\author[add1]{Marina Andretta}
		\ead{andretta@icmc.usp.br}

		\address[add1]{Department of Applied Mathematics and Statistics, Institute of Mathematical and Computer Sciences, University of São Paulo, Avenida Trabalhador São-carlense, 400, Centro, 13566-590, São Carlos, SP, Brazil.}
\end{frontmatter}
\clearpage
%\begin{frontmatter}
%		\begin{abstract}

\section*{Abstract}
			%% Text of abstract
			Planar Maximum Covering Location by Ellipses is an optimization problem where one wants to place fixed shape ellipses on the plane to cover demand points
			maximizing a function depending on the value of covered points.
			We propose new exact algorithms for two versions of this problem, one where the ellipses have to be parallel to the coordinate axis, and another where they can be freely rotated. 
			Besides finding optimal solutions for previously published instances, including the ones where no optimal solution was known, both algorithms proposed by us were able to obtain optimal solutions for some new larger instances 	with up to seven hundred demand points and five ellipses.
\subsection*{Keywords}
			Combinatorial optimization, planar maximal covering location problem, planar covering by ellipses, exact algorithms.
			%% keywords here, in the form: keyword \sep keyword

@@

%		\end{keyword}

%	\end{frontmatter}

@@

\clearpage
\section*{References}
	%% New version of the num-names style
	\bibliographystyle{apa}
	\bibliography{../references}

### 2020-03-28 defesa/article/e3p.tex
wip

In this section, we describe an experiment we made to choose a precision constant for comparing if a root of $f$ returned as an eigenvalue of its companion matrix is in the unit circle. The implementation was coded in C++, and LAPACK's ZGEEV was utilized to obtain the eigenvalues of the companion matrix of $f$ (more information about the implementation is given in \autoref{section:implementation}). For the experiment, we defined $K\in\R$, $K>0$, and considered instances with the ellipse's shape parameters $(K, \frac{K}{2})$, for $K\in\{10^j\colon j=0,\dots,10\}$.

>>>>

In this section, we describe an experiment we made to choose a precision constant for comparing if a root of $f$ returned as an eigenvalue of its companion matrix is in the unit circle. The implementation was coded in C++, and LAPACK's ZGEEV was utilized to obtain the eigenvalues of the companion matrix of $f$ (more information about the implementation is given in Section~\ref{section:implementation}). For the experiment, we defined $K\in\R$, $K>0$, and considered instances with the ellipse's shape parameters $(K, \frac{K}{2})$, for $K\in\{10^j\colon j=0,\dots,10\}$.

### 2020-03-28 defesa/article/mce.tex
wip

With all this in hand, we define \autoref{algoritmo:mce}, which goes through every possible combination in the CLS of each ellipse. As evaluating each solution can be done in $\bigO(nm)$, we have that \autoref{algoritmo:mce} has $\bigO(mn^{2m+1})$ runtime complexity. 
In \autoref{section:improvements}, we give some suggestions of improvements that can be applied to the implementation of \autoref{algoritmo:mce}, and in \autoref{section:numerical}, we analyze the results obtained by \autoref{algoritmo:mce} for some numerical experiments.

>>>>

With all this in hand, we define \autoref{algoritmo:mce}, which goes through every possible combination in the CLS of each ellipse. As evaluating each solution can be done in $\bigO(nm)$, we have that \autoref{algoritmo:mce} has $\bigO(mn^{2m+1})$ runtime complexity. 
In Section~\ref{section:improvements}, we give some suggestions of improvements that can be applied to the implementation of \autoref{algoritmo:mce}, and in Section~\ref{section:numerical}, we analyze the results obtained by Section~\ref{algoritmo:mce} for some numerical experiments.

### 2020-03-28 defesa/article/mcer.tex
wip

Following that, we introduce a lemma that is responsible for connecting the developments of this chapter with the results of \autoref{section:e3p}.
This lemma makes it possible to describe a type of solution which, for sure, is part of the equivalence class of any optimal solution.
It states that, for any ellipse that covers more than two points in a given optimal solution, an equivalent solution exists with at least one of the two properties:
\begin{itemize}
	\item The ellipse contains at least three points.
	\item The ellipse contains two points for any feasible angle.
\end{itemize}

@@

To compute $S_j^{(3)}$, we have to call the algorithm described in \autoref{section:e3p} to determine every solution of E3P for every triplet of points in $\Pp$. Even though that algorithm is $\bigO(1)$, it has a high constant factor, thus skipping it, in practice, is a good suggestion.
Given three points and an ellipse with shape parameters $(a, b)$, the following two conditions are sufficient for E3P to have no solutions, and therefore, if any of them is true, we can skip calling the algorithm to determine every solution of E3P for that instance:
\begin{itemize}
	\item The maximum distance between any of the points is greater than $2a$;
	\item The triangle's area with vertices on these three points have area greater than $\frac{3\sqrt{3}}{4}\pi ab$, which can be proved to be the greatest area of an inscribed triangle in an ellipse with shape parameters $(a, b)$.
\end{itemize}

>>>>

Following that, we introduce a lemma that is responsible for connecting the developments of this chapter with the results of Section~\ref{section:e3p}.
This lemma makes it possible to describe a type of solution which, for sure, is part of the equivalence class of any optimal solution.
It states that, for any ellipse that covers more than two points in a given optimal solution, an equivalent solution exists with at least one of the two properties:
\begin{itemize}
	\item The ellipse contains at least three points.
	\item The ellipse contains two points for any feasible angle.
\end{itemize}

@@

To compute $S_j^{(3)}$, we have to call the algorithm described in Section~\ref{section:e3p} to determine every solution of E3P for every triplet of points in $\Pp$. Even though that algorithm is $\bigO(1)$, it has a high constant factor, thus skipping it, in practice, is a good suggestion.
Given three points and an ellipse with shape parameters $(a, b)$, the following two conditions are sufficient for E3P to have no solutions, and therefore, if any of them is true, we can skip calling the algorithm to determine every solution of E3P for that instance:
\begin{itemize}
	\item The maximum distance between any of the points is greater than $2a$;
	\item The triangle's area with vertices on these three points have area greater than $\frac{3\sqrt{3}}{4}\pi ab$, which can be proved to be the greatest area of an inscribed triangle in an ellipse with shape parameters $(a, b)$.
\end{itemize}
